Studying public transit API query logs\\ to get an indication of travel flows

### 2015-12-22 paper.tex
Added histogram as an example
Update after confcall

\maketitle
\begin{abstract}
% Context
%Public transit schedules are made available on the Web in various ways, such as in a GTFS file, a route planning API, or through Linked Connections.
The iRail project hosts a route planning API for the Belgian railway company.
This webservice provides an answer to a route planning question with parameters such as ``departuretime'', ``from'' and ``to''.
% Need
In the field of urban planning, researchers need an indication of how people move between cities. 
%Yet, getting this data from official sources has proven to be troublesome.
% Task
In this paper, we analyse the query logs of the iRail API, to see whether it can tell us something about travel flows between cities in Belgium.
% Object
We studied queries for the year 2013, which contains an average of $\sim3000$ interesting queries per day ($\sim0.4\%$ of the amount of passengers per weekday).
% Findings
We illustrate \todo{X} interesting patterns that correspond to reality and we show an outlier in the data.
% Conclusion
Visualizing the data shows the query logs are a better representation than random data, yet more research is needed to understand how representative for travel flows they may be.
% Perspectives
In earlier work, we suggested the Linked Data Fragments axis to discuss new approaches to publish queryable public transit data.
Now, we suggest a discussion for gathering query logs when using Linked Data Fragments approaches.

@@

@@

@@

There are obvious caveats associated with the use of such data as proxy for actual statistical counts. 
For instance, the data only captures an intention of a passenger: it's unsure whether the person actually took the train.
Furthermore, different requests could be done by one person with an intention to travel.
On top of that, our sample used for this paper can only represent a maximum of 0.4\% of the daily traffic flow.

@@

In order to show that our data is better than random, we show that the data shows results that correspond with Belgian reality, whereas a random result would not.
To do so, we chose to assemble the smaller stations under their lager administrative unit: provinces.
The bigger cities, such as Ghent, Antwerp, Brussels, Mons and Liège, we kept.
This way, we are able to create pair of chord diagrams.

The random dataset was created as follows: \todo{...}

@@

\begin{figure}
\centering
\includegraphics[width=8.1cm,align=center]{brussels}
\caption{Pair of chord visualization of the city of Brussels: transit flows towards Brussels from elsewhere in Belgium and vice versa. Brussels acts as the central hub in the star-shaped railway network of Belgium.}
\label{fig:brussels}
\end{figure}

@@

%We can also observe the effect of "density" in Flanders where people move from the rural towards cities. Look at weekday-mornings: commuting towards Gent, Brussels, Antwerp. This is not as obvious in the Walloon region (except for Liege). The reverse occurs on weekday-evenings. Local patterns can also be observed and this corresponds to what we know: "cities are the central hubs".
%Walloon region: everything converges on Brussels
%I found that whereas there's almost no movement from Antwerp city in the week to West Flanders, there is movement from Antwerp in the weekend to the coast and back. 
\begin{figure}
\centering
\includegraphics[width=8.1cm,align=center]{antwerp}
\caption{Pair of chord visualization of the city of Antwerp: transit flows towards Antwerp from elsewhere in Belgium and vice versa. The biggest transit flows to elsewhere in the province of Antwerp Central station is the hub for the province.}
\label{fig:antwerp}
\end{figure}

@@

Thoughts on publishing transport data on the web and query logs in the case of GTFS, Linked Connections\cite{lc} (linked data fragments\cite{ldf}) and webservices.

@@

Query logs are very interesting

>>>>

\maketitle
\begin{abstract}
% Context
%Public transit schedules are made available on the Web in various ways, such as in a GTFS file, a route planning API, or through Linked Connections.
The iRail project hosts a route planning API for the Belgian railway company.
This webservice provides an answer to a route planning question with parameters such as ``departuretime'', ``from'' and ``to''.
% Need
In the field of urban planning, researchers need an indication of how people move between cities. 
%Yet, getting this data from official sources has proven to be troublesome.
% Task
In this paper, we analyse the query logs of the iRail API, to see whether it can tell us something about travel flows between cities in Belgium.
% Object
We studied queries for the year 2013, which contains an average of $\sim3000$ interesting queries per day ($\sim0.4\%$ of the amount of passengers per weekday).
% Findings
We illustrate \todo{X} interesting patterns that correspond to reality and we show an outlier in the data.
% Conclusion
We could not find a way \todo{to validate \ldots}, yet more research is needed to understand how representative for travel flows they may be.
% Perspectives
In earlier work, we suggested the Linked Data Fragments axis to discuss new approaches to publish queryable public transit data.
Now, we suggest a discussion for gathering query logs when using Linked Data Fragments approaches.

@@

\todo{Rework introduction at the end}

@@

\todo{Write relwork at the end}

@@

As we are dealing with query logs, privacy becomes an issue~\cite{silvestri}.
However, we do not publish the IP addresses with each query, which makes it hard to track single users.
To remove all doubt, we have asked and received the consent of the privacy commission of Belgium.

@@

\subsection{Processing}
\todo{Each data record contains the names of both origin and destination (OD) stations, the user agent that made the query todo: Pieter this needs definition, as well as the time when the query was made.

As we were specifically interested in examining morning rush hour travel on weekdays, we filtered the dataset to exclude data records created outside the range of xxx hrs and xxx hrs on weekdays, and all data records created over the weekends.

The iRail API serves a wide range of  applications that vary in terms of software architecture and functionality. End user applications such as BeTrains and Railer Pieter xxx and xxx query for time schedules when explicitly directed by the user. On the other hand, automated services like xxx make routine queries to gather data for further computation. Naturally, such automated queries are detected based on xxx.

We exclude all automated queries to refine the dataset and remove potential bias.

Queries made by automated user agents like xxx and xxx were also excluded from the dataset to remove biases that may over represent connections between certain stations.

Additionally, data records with unparsable origin or destination station names due to spelling mistakes or invalid character encoding were also removed from the data set.}

\subsection{Visualization}

\todo{Alvin}

@@

\begin{figure}
\centering
\includegraphics[width=8.1cm,align=center]{histogram}
\caption{Histogram of one weekday}
\label{fig:brussels}
\end{figure}

In this chapter, we report on the results of studying visualizations with experts of Trein Tram Bus, a not for profit passenger federation.

\subsection{Commuting pattern}

We see people leaving in the morning and return in the evening (see \cref{fig:brussels} and todo)

@@

\subsection{Flanders is polycentric, Wallonia monocentric}

We can also observe the effect of "density" in Flanders where people move from the rural towards cities. Look at weekday-mornings: commuting towards Gent, Brussels, Antwerp. This is not as obvious in the Walloon region (except for Liege). The reverse occurs on weekday-evenings. Local patterns can also be observed and this corresponds to what we know: ``cities are the central hubs''.
Walloon region: everything converges on Brussels

\subsection{Inexplicable results}

We also see weird things happening in Antwerp, where a large amount of people go to the province in the morning on weekdays, while the oposite happens in Liege.
When looking into the data, the Antwerp pattern appears due to one specific route which is queried each morning.
Is there a group of colleagues all using the same app taking the same train daily?
Is it someone who checks this trip over and over again for delays?
We can only guess\ldots

\begin{figure}
\centering
\includegraphics[width=8.1cm,align=center]{brussels}
\caption{Pair of chord visualization of the city of Brussels: transit flows towards Brussels from elsewhere in Belgium and vice versa.}
\label{fig:brussels}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=8.1cm,align=center]{antwerp}
\caption{Pair of chord visualization of the city of Antwerp: transit flows towards Antwerp from elsewhere in Belgium and vice versa.}
\label{fig:antwerp}
\end{figure}

@@

%\subsection{Intermodal route planning}

@@

The contribution of this paper is twofold.
For the first time, we studied query logs to find travel patterns.
Due to lack of information we could not evaluate our results, yet the visualizations look promising...
There are obvious caveats associated with the use of such data as proxy for actual statistical counts. 
For instance, the data only captures an intention of a passenger: it's unsure whether the person actually took the train.
Furthermore, different requests could be done by one person with an intention to travel.
On top of that, our sample used for this paper can only represent a maximum of 0.4\% of the daily traffic flow.

The second contribution is our position that we suggest to decouple the query logs from the query execution.
